---
title: "R Notebook"
output: html_notebook
---

# Intial Setup

```{r}
#setup working directory
setwd("C:/Users/natha/Dropbox/R_training")

#load libraries
library(tidyverse)
library(ggthemes)
df <- read_csv("housing.csv")

```

Overview
Layers in Building a Plot
1. Data 
2. Aesthetics - Scales which data are mapped on 
3. Geometries - Shapes used to represent data
4. Facets - Splitting by rows and columns
5. Statistics - Models and summaries
6. Coordinates - Space used for plotting
7. Theme - Design elements

# General Format
```{r}

#basics
df %>%
  ggplot(aes()) +
  geom()

#labels
+ labs (title = "Nate's Test Plot",
subtitle = "Testing formatting",
x = "Date"
y = "Price ($M)",
color = "Region"
)

#themes
+ theme_fivethirtyeight()
+ theme(axis.title = element_text())    --add the axis titles back in

#emphasize Data
+ geom_line (size = 1.5, alpha = 0.8)
+ scale_linetype_manual (values = c("dashed", "solid"), guide = "none")

#use a set of colors from a package
+ scale_color_brewer (palette = "Set1")

#Manual determine colors
myColors = c("#A5511A", #DFC27D")

+ scaled_color_manual (values = myColors)
+ theme(text = element_text(family = "Rubik"))

```

# One Variable (continuous)

```{r}
# histogram
df %>%
  ggplot(aes(median_income)) + 
  geom_histogram()

# density
df %>%
  ggplot(aes(median_income)) + 
  geom_density()

df %>%
  ggplot(aes(median_income)) + 
  geom_dotplot()

df %>%
  ggplot(aes(median_income)) + 
  geom_freqpoly()

df %>%
  ggplot(aes(sample=median_income)) + 
  geom_qq()

df %>%
  ggplot(aes(ocean_proximity)) + 
  geom_bar()


```

# Two Variables (both continuous)

```{r}
# histogram
df %>%
  ggplot(aes(total_rooms,total_bedrooms)) + 
  geom_point()

df %>%
  ggplot(aes(total_rooms,total_bedrooms)) + 
  geom_quantile()

df %>%
  ggplot(aes(total_rooms,total_bedrooms)) + 
  geom_rug()

df %>%
  ggplot(aes(total_rooms,total_bedrooms)) + 
  geom_smooth()

df %>%
  ggplot(aes(total_rooms,total_bedrooms)) + 
  geom_bin2d()

df %>%
  ggplot(aes(total_rooms,total_bedrooms)) + 
  geom_density_2d()

df %>%
  ggplot(aes(total_rooms,total_bedrooms)) + 
  geom_hex()

df %>%
  ggplot(aes(total_rooms,total_bedrooms)) + 
  geom_area()

df %>%
  ggplot(aes(total_rooms,total_bedrooms)) + 
  geom_line()

df %>%
  ggplot(aes(total_rooms,total_bedrooms)) + 
  geom_step()

```

# Two Variables (one discrete, one continuous)

```{r}
# histogram
df %>%
  ggplot(aes(ocean_proximity,total_bedrooms)) + 
  geom_col()

df %>%
  ggplot(aes(ocean_proximity,total_bedrooms)) + 
  geom_boxplot()

df %>%
  ggplot(aes(ocean_proximity,total_bedrooms)) + 
  geom_dotplot(binaxis = "y", stackdir = "center")

df %>%
  ggplot(aes(ocean_proximity,total_bedrooms)) + 
  geom_violin()

```

# Two Variables (two discrete)

```{r}
# histogram
df$ocean_proximity <- as.factor(df$ocean_proximity)
df$median_income <- as.factor(df$median_income)

df %>%
  ggplot(aes(ocean_proximity,median_income)) + 
  geom_count()

df %>%
  ggplot(aes(ocean_proximity,median_income)) + 
  geom_jitter()

# scatterplot
df %>%
  ggplot(mapping = aes(x = total_rooms, y = total_bedrooms,color=ocean_proximity)) + 
  geom_point() +
  geom_smooth(method = "lm")

# line graph
df %>%
  ggplot(mapping = aes(x = total_rooms, y = median_house_value)) + 
  geom_line()

# bar chart
df %>%
  ggplot(mapping = aes(x = ocean_proximity, fill = ocean_proximity)) + 
  geom_bar()+ 
  scale_fill_discrete()

#labels on bars example
geom_text(aes(label = freq, y = freq+10), size = 10)

df %>% 
  group_by(ocean_proximity) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  mutate(prox = as.factor(ocean_proximity), 
         prox = fct_inorder(ocean_proximity)) %>%
  ggplot(aes(x = prox,
             y = n,
             fill = prox)) +
  geom_bar(stat = "identity",
           show.legend = F,
           color = "black") +
  geom_text(aes(label = n, y = n + 50, size = 2
  ))





```

Geom

# Graphical primitives
geom_blank(): display nothing. Most useful for adjusting axes limits using data
geom_point(): points
geom_path(): paths
geom_ribbon(): ribbons, a path with vertical thickness
geom_segment(): a line segment, specified by start and end position
geom_rect(): rectangles
geom_polygon(): filled polygons
geom_text(): text

# One variable
## Discrete
geom_bar(): display distribution of discrete variable

## Continuous
geom_histogram(): bin and count continuous variable, display with bars
geom_density(): smoothed density estimate
geom_dotplot(): stack individual points into a dot plot
geom_freqpoly(): bin and count continuous variable, display with lines

# Two variables
## Both continuous
geom_point(): scatterplot
geom_quantile(): smoothed quantile regression
geom_rug(): marginal rug plots
geom_smooth(): smoothed line of best fit
geom_text(): text labels
## Show distribution
geom_bin2d(): bin into rectangles and count
geom_density2d(): smoothed 2d density estimate
geom_hex(): bin into hexagons and count
## At least one discrete
geom_count(): count number of point at distinct locations
geom_jitter(): randomly jitter overlapping points
## One continuous, one discrete
geom_bar(stat = "identity"): a bar chart of precomputed summaries
geom_boxplot(): boxplots
geom_violin(): show density of values in each group
## One time, one continuous
geom_area(): area plot
geom_line(): line plot
geom_step(): step plot
# Display uncertainty
geom_crossbar(): vertical bar with center
geom_errorbar(): error bars
geom_linerange(): vertical line
geom_pointrange(): vertical line with center
# Spatial
geom_map(): fast version of geom_polygon() for map data
# Three variables
geom_contour(): contours
geom_tile(): tile the plane with rectangles
geom_raster(): fast version of geom_tile() for equal sized tiles

```{r}
dplot <- ggplot(diamonds, aes(color, fill = cut)) + 
  xlab(NULL) + ylab(NULL) + theme(legend.position = "none")
# position stack is the default for bars, so `geom_bar()` 
# is equivalent to `geom_bar(position = "stack")`.
dplot + geom_bar()
dplot + geom_bar(position = "fill")
dplot + geom_bar(position = "dodge")
dplot + geom_bar(position = "identity", alpha = 1 / 2, colour = "grey50")

ggplot(diamonds, aes(color, colour = cut)) + 
  geom_line(aes(group = cut), stat = "count") + 
  xlab(NULL) + ylab(NULL) + 
  theme(legend.position = "none")

facet_wrap(~variable)
facet_grid() lays out plots in a 2d grid, as defined by a formula: 
  base + facet_grid(. ~ cyl)
base + facet_grid(drv ~ .)
base + facet_grid(drv ~ cyl)


```

# Themes 

https://www.cedricscherer.com/2019/08/05/a-ggplot2-tutorial-for-beautiful-plotting-in-r/
http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf

```{r}
# set a theme
theme_set(theme_bw())

# 
color = "firebrick"
shape = "diamond"
size = 2

# Add axes
labs (x = "Year", y = "Temperature (째F)")
xlab("Year") +
  ylab("Temperature (째F)")

axis.title.x = element_text(vjust = 0, size = 15)
axis.title.y = element_text(vjust = 2, size = 15)

theme(axis.title = element_text(size = 15, color ="firebrick",face = "italic"))

ggtitle("Temperatures in Chicago")

title = "Temperatures in Chicago",
subtitle = "Seasonal pattern of daily temperatures from 1997 to 2001",

theme(plot.title = element_text(family = "Bangers", hjust = .5, size = 25))

theme(legend.position = "top")

guides(color = guide_legend(override.aes = list(size = 6)))

gb <- ggplot(chic, aes(x = date, y = temp, color = temp)) +
  geom_point() +
  labs(x = "Year", y = "Temperature (째F)", color = "Temperature (째F):")

gb + scale_color_continuous() 

gb + scale_color_gradient()

gb + scale_color_gradient(low = "darkkhaki",
                          high = "darkgreen")

gb + scale_color_gradient2(midpoint = mid, low = "#dd8a0b",
                           mid = "grey92", high = "#32a676")

geom_vline(xintercept = 2, linetype = "dashed", size = 1)


```


# Lines
Different options using integer or name

```{r}
lty <- c("solid", "dashed", "dotted", "dotdash", "longdash", "twodash")
linetypes <- data.frame(
  y = seq_along(lty),
  lty = lty
) 
ggplot(linetypes, aes(0, y)) + 
  geom_segment(aes(xend = 5, yend = y, linetype = lty)) + 
  scale_linetype_identity() + 
  geom_text(aes(label = lty), hjust = 0, nudge_y = 0.2) +
  scale_x_continuous(NULL, breaks = NULL) + 
  scale_y_reverse(NULL, breaks = NULL)

```





